Model Search in Relational Domains

### 2016-10-22 deductions.tex
part of deductions

To make our algorithm efficient, we have to focus abstract
transformers on performing only the minimally necessary
work. 
%
We thus define a specialised variant of the abstract transformer to compute
deductions w.r.t.\ a given subdomain $\subdomain\subseteq \domain$.
A subdomain contains a chosen subset of the elements in $\domain$ including $\bot$ and $\top$
\begin{equation}\label{eq:at2}
\abstrans{\domain}{\constraint}^\subdomain(\absval)=\absval\meet_\domain \alpha_\subdomain(\{\val\mid \val\in\gamma_\domain(\absval), \val\models \constraint\})
\end{equation}
For $\subdomain=\domain$, the abstract transformer behaves in the same way as the one defined in Section~\ref{sec:domains}.

@@

The second advantage of Eq.~(\ref{eq:at2}) is that we can compute the
\emph{closure} operation for relational domains in a lazy manner. To
this end, we construct a $\subdomain=\makesubdomain_\domain(\vars)$
which is sufficient to perform one step of the closure operation each
time the abstract transformer is applied.
%
For example, let us consider $\domain=\octagons[\{x,y,z\}]$ and
$\vars=\{y\}$. \rmcmt{A binary} octagon \rmcmt{element} relate at 
most two variables. Thus it is sufficient to consider the subdomain
$\makesubdomain_\domain(\{y\})=\octagons[\{y\}]\cup\octagons[\{x,y\}]\cup\octagons[\{y,z\}]$,
which will compute the one-step transitive relations of $y$ with each
of the other variables. If this indeed affects $x$ and $z$ then their
relations will be inferred in a later step of the propagation
algorithm, thus gradually computing the closure. Note that this might
increase the number of propagation steps, but it reduces the size of
the domain used in the abstract transformer.

@@

Algorithm~\ref{Alg:ms} presents the deduction phase $\deduce$ in 
abstract model search procedure.  The input to the algorithm is 
the set of abstract transformers, a propagation trail and a reason 
trail.  Additionally, the procedure $\deduce$ is parameterised with 
a propagation heuristic.  The propagation heuristics provides two 
functions $\initworklist$ and $\updateworklist$.  Given a set of 
variables in the \rmcmt{constraint system}, a subdomain is essentially a 
subset of variables \rmcmt{how we are defining subdomain?} that guides the propagations in forward, 
backward or multi-way.  For example, the subdomain consist of variables 
in lhs in forward propagation heuristic.  For backward, the subdomain 
consist of variables in rhs, whereas the subdomain contains all variables 
for multi-way propagation.  A {\em worklist} is a queue which contains 
a set of abstract deduction transformers.  
The initWorklist initializes the worklist 
following a heuristics choice for propagation strategies.  A {\em forward} 
propagation strategy initializes the worklist with abstract deduction 
transformers that contains constants in the right-hand side. Whereas, a 
{\em backward} iteration strategy starts with a set of assertions. On the 
other hand, a {\em multi-way} iteration strategy initializes the worklist 
with set of all transformers and performs both forward and backward 
propagations in tandem.  \rmcmt{wrong to say abstract deduction transformer that 
has lhs, rhs since since it also includes the live variables}.

Given a constraint system and a domain, an updateWorklist function adds 
new abstract deduction transformers to the worklist for those variables in 
the subdomain that appear in newDeductions.  It also updates the subdomain of 
existing abstract deduction transformers that are in the worklist.

The function
$\onlynew(\absval)=\bigsqcap(\decomp(\absval)\setminus\decomp(\abs(\trail)))$
filters out all known meet irreducibles in the abstract value.

For every abstract deduction transformer in the worklist, 
the set of new meet irreducibles inferred by the domain is added 
to the $\trail$.  Additionally, the reason trail $\reasons$ is 
updated with the corresponding abstract transformer as the reason for inferring 
those new meet irreducibles.  If a conflict is deduced, 
%that is, $\mathit{meet\_irrd} = \bot$ is inferred by the domain, 
the algorithm terminates with \textsf{conflict}.  However,
when a fixed-point is reached, the abstract value $v$ computed from 
the trail is checked for set of models.  This is determined by 
checking whether the set of assignments in $v$ is 
$\gamma$-complete~\cite{dhk2013-popl}. If $v$ is $\gamma$-complete, that is, the 
abstract valuation $v$ is abstractly satisfying, the abstract model search 
terminates with \textsf{sat}.  Otherwise, the algorithm returns \textsf{unknown} and the 
model search procedure makes a new decision.

>>>>

To make our algorithm efficient, we have to focus abstract
transformers on performing only the minimally necessary
work. 
%
We thus define a specialised variant of the abstract transformer to compute
deductions w.r.t.\ a given subdomain $\subdomain\subseteq \domain$,
which we call \emph{abstract deduction transformer}.
%
A subdomain contains a chosen subset of the elements in $\domain$ including $\bot$ and $\top$
\begin{equation}\label{eq:at2}
\abstrans{\domain}{\constraint}^\subdomain(\absval)=\absval\meet_\domain \alpha_\subdomain(\{\val\mid \val\in\gamma_\domain(\absval), \val\models \constraint\})
\end{equation}
For $\subdomain=\domain$, the abstract deduction transformer behaves in the same way as the abstract transformer defined in Eq.~\ref{eq:abstrans} in Section~\ref{sec:domains}.

@@

The second advantage of Eq.~(\ref{eq:at2}) is that we can compute the
\emph{closure} operation for relational domains in a lazy manner. To
this end, we construct a $\subdomain=\makesubdomain_\domain(\vars)$
which is sufficient to perform one step of the closure operation each
time the abstract transformer is applied.
%
For example, let us consider $\domain=\octagons[\{x,y,z\}]$ and
$\vars=\{y\}$. An octagonal inequality relates at 
most two variables. Thus it is sufficient to consider the subdomain
$\makesubdomain_\domain(\{y\})=\octagons[\{y\}]\cup\octagons[\{x,y\}]\cup\octagons[\{y,z\}]$,
which will compute the one-step transitive relations of~$y$ with each
of the other variables. 
%
\Omit{
%TODO
Hence, we do not compute the full closure in the full 
domain $\octagons[\{x,y,z\}]$, but we compute only a single step of the closure when the abstract deduction transformer is applied.
%
Only if $\abstrans^\subdomain$ deduces new information about $x$ or $z$
then the next step of the closure will be 
}
If this indeed affects $x$ and~$z$ then their
relations will be inferred in a later step of the propagation
algorithm, thus gradually computing the closure.

%Note that this might
%increase the number of propagation steps, but it reduces the size of
%the subdomain used in the abstract deduction transformer.

@@

Algorithm~\ref{Alg:ms} presents the deduction phase $\deduce$ in 
abstract model search procedure.  The input to the algorithm is 
the set of abstract transformers, a propagation trail and a reason 
trail.  Additionally, the procedure $\deduce$ is parameterised with 
a propagation heuristic. 
%
The algorithm maintains a  {\em worklist}, which is a queue that contains 
abstract deduction transformers.  
The propagation heuristics provides two 
functions $\initworklist$ and $\updateworklist$.
The order of the elements in the worklist and the subdomains $\subdomain$
attached to each of the abstract deduction transformers
determine the propagation strategy (forward, backward, multi-way).
%
The {\em forward} propagation strategy initialises the worklist with
abstract deduction transformers that contains constants in the
right-hand side and the subdomain is constructed via $\makesubdomain$
from variables in left-hand side of the equality constraints
originating from the assignment statements in the program.
%
The {\em backward} iteration strategy intialises the worklist with the assertions and the subdomain is constructed from the right-hand side variables.
%
The {\em multi-way} iteration strategy initialises the worklist 
with set of all transformers and each subdomain contains the variables occurring in the respective transformers.

Depending on the propagation heuristics $\updateworklist$ adds
abstract deduction transformers to the worklist that contain variables
that appear in $\newdeductions$.  It also updates the subdomains of
existing abstract deduction transformers that are in the worklist.

The function
$\onlynew(\absval)=\bigsqcap(\decomp(\absval)\setminus\decomp(\abs(\trail)))$
filters out all meet irreducibles that are already on the trail.

\Omit{
For every abstract deduction transformer in the worklist, 
the set of new meet irreducibles inferred by the domain is added 
to the $\trail$.  Additionally, the reason trail $\reasons$ is 
updated with the corresponding abstract transformer as the reason for inferring 
those new meet irreducibles.  If a conflict is deduced, 
%that is, $\mathit{meet\_irrd} = \bot$ is inferred by the domain, 
the algorithm terminates with \textsf{conflict}.  However,
when a fixed-point is reached, the abstract value $v$ computed from 
the trail is checked for set of models.  This is determined by 
checking whether the set of assignments in $v$ is 
$\gamma$-complete~\cite{dhk2013-popl}. If $v$ is $\gamma$-complete, that is, the 
abstract valuation $v$ is abstractly satisfying, the abstract model search 
terminates with \textsf{sat}.  Otherwise, the algorithm returns \textsf{unknown} and the 
model search procedure makes a new decision.    
}
